Leadership across cultures and contexts: an exploratory analysis

### 2020-05-16 leadership_across_cultures_contexts.Rmd
Adding text to SI.

Below we briefly review leadership functions and qualities in WEIRD contexts, to suppliment our discussions of non-WEIRD leaderhsip dynamics in the main text.

### WEIRD leadership functions

<!-- WEIRD ORGNIZATIONAL STRUCTURE TELLS FUNCTIONS -->
In WEIRD societies, community and political leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990]. Other public service, nonprofit leaders and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders, to name a few [@martiskainen_role_2017; @azzam_community_2003; @grace_school_2005]. These positions typically are associated with institutionalized responsibilities and specific functions. The structure of such organizations are generally hierarchical with nested bodies and an internal 'chain of command.' At higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization [@keohane_thinking_2010]. They are not typically directly involved in resolving interpersonal conflicts between individuals, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications. Within local communities and in smaller-scale economic groups, however, leaders in WEIRD populations *are* likely to be direlty involved in mediating conflicts and managing local cooperative projects [e.g., @acheson1988lobster; @rich_dynamics_1980].

@@

### WEIRD leadership qualities

Drawing on cross-national data through the GLOBE project from over 15,000 middle managers in businesses across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership. In particular, inducing motivation, having foresight, being encouraging and communicative, trustworthiness, and building self-confidence among followers are suggested to be universal attributes of outstanding leadership in managerial contexts.

@@

### WEIRD costs and benefits

<!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - LSS -->
Conceptualizing leadership as a possible solution to problems associated with collective goods and actions [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2019], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation. In this framework, leaders and followers voluntary participate in a cooperative enterprise with the expectation that benefits will outweigh the associated costs. @rich_dynamics_1980 adopts this framework of collective costs and benefits to understand leadership and group dynamics in WEIRD urban neighborhood organizations, which can be viewed as self-interested groups that face group-wide problems and necessarily share public benefits. The costs associated with leadership in neighborhood organizations follow those associated with other collective enterprises (e.g., investments of time and energy in management and maintenance of group coordination). @rich_dynamics_1980 conceptualizes four categories of the benefits associated with neighborhood organization. Leaders of neighborhood organizations (1) may receive benefits of the public goods produces from collective actions (as other non-leaders would as well); (2), may receive selective benefits from surpluses produced through collective actions; (3) may receive incidental benefits from their leadership role; and (4) leaders of neighborhood organizations may enjoy their service in the position. @rich_dynamics_1980 interviewed 54 officers from 11 metropolitan neighborhood organizations in the Midwest US. Results from extensive interviews suggest personal fulfillment and enjoyment are the main motivational drivers of participation in leadership roles of neighborhood organizations, suggesting the process of leadership for individuals is distinct from theoretical models rooted in economics. Therefore, a primary benefit of leadership, in some contexts, is likely to be intrinsic personal fulfillment and the opportunity to execute personal values of community service.

@@

@@

Leadership dimensions for which leader sex significantly improved fit. Note that very few text records (N=`r sum(all_data$demo_sex=='both')`) explicitly mentioned both male and female leaders. Hence, although not depicted here, the confidence intervals for the estimated marginal means for this level are wide.

@@

@@

>>>>

Below we briefly review leadership functions and qualities in WEIRD contexts to suppliment our discussions of non-WEIRD leaderhsip dynamics in the main text.

**WEIRD leadership functions**. In WEIRD societies, community and political leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990]. Other public service, nonprofit leaders and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders, to name a few [@martiskainen_role_2017; @azzam_community_2003; @grace_school_2005]. These positions typically are associated with institutionalized responsibilities and specific functions. The structure of such organizations are generally hierarchical with nested bodies and an internal 'chain of command.' At higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization [@keohane_thinking_2010]. They are not typically directly involved in resolving interpersonal conflicts between individuals, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications. Within local communities and in smaller-scale economic groups, however, leaders in WEIRD populations *are* likely to be direlty involved in mediating conflicts and managing local cooperative projects [e.g., @acheson1988lobster; @rich_dynamics_1980].

@@

**WEIRD leadership qualities**. Drawing on cross-national data through the GLOBE project from over 15,000 middle managers in businesses across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership. In particular, inducing motivation, having foresight, being encouraging and communicative, trustworthiness, and building self-confidence among followers are suggested to be universal attributes of outstanding leadership in managerial contexts.

@@

**WEIRD leadership costs and benefits**. Conceptualizing leadership as a possible solution to problems associated with collective goods and actions [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2019], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation. In this framework, leaders and followers voluntary participate in a cooperative enterprise with the expectation that benefits will outweigh the associated costs. @rich_dynamics_1980 adopts this framework of collective costs and benefits to understand leadership and group dynamics in WEIRD urban neighborhood organizations, which can be viewed as self-interested groups that face group-wide problems and necessarily share public benefits. The costs associated with leadership in neighborhood organizations follow those associated with other collective enterprises (e.g., investments of time and energy in management and maintenance of group coordination). @rich_dynamics_1980 conceptualizes four categories of the benefits associated with neighborhood organization. Leaders of neighborhood organizations (1) may receive benefits of the public goods produces from collective actions (as other non-leaders would as well); (2), may receive selective benefits from surpluses produced through collective actions; (3) may receive incidental benefits from their leadership role; and (4) leaders of neighborhood organizations may enjoy their service in the position. @rich_dynamics_1980 interviewed 54 officers from 11 metropolitan neighborhood organizations in the Midwest US. Results from extensive interviews suggest personal fulfillment and enjoyment are the main motivational drivers of participation in leadership roles of neighborhood organizations, suggesting the process of leadership for individuals is distinct from theoretical models rooted in economics. Therefore, a primary benefit of leadership, in some contexts, is likely to be intrinsic personal fulfillment and the opportunity to execute personal values of community service.

@@

Below we plot the estimated margianl means of the effects for level of evidence for each of the `r vary_n` (out of `r nrow(multi_allvars)` total) which produced a model that showed an improvement *with* the four factor predictors over an intercepet only model, by the variable which produced the model improvement. See Figure \@ref(fig:heatmapdrop1)A in the main text. Figure \@ref(fig:heatmapgroups) plots the ledaerhsip dimensions which varied by group structure, Figure \@ref(fig:heatmapsubsis) plots the leadership dimensions which varied by subsistence stragegy, Figure \@ref(fig:heatmapregion) plots the leadership dimnenbsions which varied by contintental region, and Figure \@ref(fig:heatmapsex) plots the leadership dimensions which variaed by leader sex. Table \@ref(tab:variablevars) displays the nbumber of text records providing evidence for each of these the `r vary_n` varying leaderhsip dimennsions and the results of logistic mexed effects models including the AIC scores for the intercept only models, AIC values for models with the four predictors, the AIC difference, and adjusted p-values for the four factor coefficents. Table \@ref(tab:universalvars) displays the same information for the remaining leaderhsip dimensions which did not substantially vary by our four factors.

@@

@@

Note that very few text records (N=`r sum(all_data$demo_sex=='both')`) explicitly mentioned both male and female leaders. Hence, although not depicted here, the confidence intervals for the estimated marginal means for this level are wide. See Table \@ref(tab:sexsubtable) for the distribution of leader sex data by subsistence strategy.

@@

We explored logistic PCA to reduce the dimensionality of the `r length(quality_vars)` leader quality variables and the `r length(function_vars)` leader function variables, all of which take binary values (0, 1) using the logisticPCA package. This algorithm projects the natural parameters of the saturated model onto the lower $k-$dimension space that minimizes the Bernoulli deviance [@logisticPCA]. Prior to analysis we removed all rows that had no evidence for any variable, which resulted in `r nrow(df_qual)` text records for the leader quality variables and `r nrow(df_func)` text records for the leader function variables. We used cross-validation to optimize rank, $k$, and a tuning parameter, $m$ (the value to approximate the saturated model), per recommendation of the package authors.
